---
title: "R Notebook"
output: html_notebook
---
# R in a Nutshell

Author: 

## Preface

[GNU foundation's FAQ on GPL](http://www.gnu.org/licenses/gplfaq) - Handy when it comes to rights to extensions written for R (when you write code that is only interpreted by R, then no rigth bounds apply)

## Part I. R Basics

- installing
- packages (installing, usage)
- quick tutorial
- overview of R features

### Chapter 1. Getting and Installing R

### Chapter 2. The R User Interface

#### Command line editing

Lets user type commands into R interactive shell and get computed results immediately

#### Batch mode 

Provides way to run large set of commands in sequence and save the results to a file using `CMD` or `RScript` command

Example:
```
$ R CMD BATCH generate_graphs.R
```
or
```
$ RScript generate_graphs.R
```
##### Executable batch file

Contents (mind the shebang)
```
#! /usr/bin/env RScript

print("Hello world!");
```
Commands
```
# making executable
$ chmod +x hello_world.R

# executing
$ ./hello_world.R
```

##### Executing batch file from inside R

`source` command

#### R and other technologies (some libs are obsolete)

- MS Excel (many packages)
- rApache (R analysis in web app)
- Rserve (binary R server for multiple users)
- ESS - Emacs Speaks Statistics (package for using R inside Emacs)

### Chapter 3. A Short R Tutorial

In R `"Hello world."` is a character vector of length 1.

`c("Hello","World")` is a character vector of length 2.

To test it type: 

```
> length(c('hello','hello'))
```

Therefore there is no string concept in R, each character vector can store more than one character.

#### Functions

Functions take arguments and processes them, they can take form of `f(...)` or operator (e.g. addition `+`, exponentiation `^` or equality `==` operator)

Functions are assigned like variables
```
f <- function(x,y) {c(x+1, y+1)}
```
(to see code for existing function, type it's name without `()`)

#### Data Structures

Array - one data type vector that's associated with a dimension attribute

  - can have more than 2 dimensions
  
```
> a <- array(c(1:12), dim=c(3,4))
```
Matrix - two dimensional array

List  

  - vector containing multiple objects of possibly varius data types
  - each component in a list can be named
  - objects can be referred by location or name
  - can contain other lists

```{r}
# a list containing two strings
e <- list(thing='hat', size='8.25')
e
e$thing
e[1]
e[[1]]

g <- list(element_1 = 'element one',other_list = e)
g
```

Data frame

  - list
  - contains multiple vectors (optionally of various types)
  - each vector has to be the same length

```{r}
# creating df
teams <- c('PHI','NYM','FLA','ATL','WSN')
w <- c(92,89,94,72,59)
l <- c(70,73,77,90,102)
nleast <- data.frame(teams,w,l)
nleast

# get number of loses by Florida Marlins
nleast$l[nleast$teams=='FLA']
```

#### Objects and Classes

Functions available only for specific classes are called methods (duh)